### 2012-08-21 doc/prelim/document/chapters/introduction/introduction.tex
adding intro and research proposal chapter

\svnidlong{$LastChangedBy$}{$LastChangedRevision$}{$LastChangedDate$}{$HeadURL: http://freevariable.com/dissertation/branches/diss-template/frontmatter/abstract.tex $}
\vcinfo{}

In many fields of engineering and physics, nonlinear problems are a
primary focus of study. Recent focus on multiple physics systems (and
in particular nuclear systems) adds a new level of complexity to
common nonlinear systems as solution strategies change when they are
coupled to other problems. Furthermore, a desire for predictive
simulations to enhance the safety and performance of engineered
systems creates a need for extremely high fidelity computations to be
performed for these coupled systems as a means to capture effects not
modeled by coarser methods. In order to achieve this high fidelity,
state-of-the-art computing facilities must be leveraged in a way that
is both efficient and considerate of hardware-related issues. As
scientific computing moves towards exascale facilities with machines
of $O(1,000,000)$ cores already coming online, new algorithms to solve
these complex problems must be developed to leverage this new
hardware. Issues such as resiliency to node failure and scaling to
large numbers of cores will be pertinent to robust algorithms aimed at
this new hardware. Considering these issues, this thesis proposes a
novel stochastic method to advance solution techniques for nonlinear
problems with a focus on coupled physics systems.

>>>>

\chapter{Introduction}
\label{ch:introduction}

In many fields of engineering and physics, linear and nonlinear
problems are a primary focus of study. Recent focus on multiple
physics systems (and in particular nuclear systems) adds a new level
of complexity to common nonlinear systems as solution strategies
change when they are coupled to other problems. Furthermore, a desire
for predictive simulations to enhance the safety and performance of
engineered systems creates a need for extremely high fidelity
computations to be performed for these coupled systems as a means to
capture effects not modeled by coarser methods.

In order to achieve this high fidelity, state-of-the-art computing
facilities must be leveraged in a way that is both efficient and
considerate of hardware-related issues. As scientific computing moves
towards exascale facilities with machines of $O(1,000,000)$ cores
already coming online, new algorithms to solve these complex problems
must be developed to leverage this new hardware. Issues such as
resiliency to node failure and scaling to large numbers of cores will
be pertinent to robust algorithms aimed at this new
hardware. Considering these issues, this thesis proposes a the
development of a massively parallel stochastic method for linear
problems and a novel stochastic method to advance solution techniques
for nonlinear problems with a focus on coupled physics systems.

This preliminary report oultines the conventional methods used in
practice for solving linear and nonlinear problems to provide a
mathematical basis upon which to build new algorithms that aim to
solve some of the aforementioned issues. These new algorithms will be
outlined in full with links made to past work and their potential for
offering improvements to the computational physics community. Progress
to date will be reported including the development of a new physics
framework that will serve as a test bed for these new methods. A
research plan will then be provided that continues the development and
implementation of these new methods and tests them in practical
settings.

\section{Motivation}
\label{sec:motivation}
